---
title: "Supplementary Document: MCMC Diagnostics"
subtitle: "A Bayesian Capture-Recapture model of vector-reservoir interaction in an ecological setting: a reservoir-targeted vaccine field study against Borrelia burgdorferi"
author: "Felix M. Pabon-Rodriguez, Grant D. Brown, Maria Gomes-Solecki, and Christine A. Petersen"
date: ''
output: pdf_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries
options(repos="https://cran.rstudio.com")
install_load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

my_packages <- c("readxl", "dplyr", "coda", "ggplot2",
                 "lattice", "bayesplot", "BayesPostEst", "ggmcmc", 
                 "RCurl", "truncnorm", "kableExtra", "mvtnorm", "rlist", 
                 "extraDistr", "msm", "tmvtnorm", "runjags", "plotrix",
                 "lubridate", "ggpubr", "stringr", "nimble", "tibble",
                 "igraph", "parallel", "doParallel", "MCMCvis", "LaplacesDemon")
invisible(install_load(my_packages))

########################################################################
# Posterior samples
########################################################################

options(digits = 4)

## 2020

DP_file2020 <- readRDS("fit_DP2020_idx103_new_official.rds")
DP_mcmc2020 <- as.mcmc.list(lapply(1:3,function(x){as.mcmc(DP_file2020[[x]]$samples)}))
combined_listDP2020 <- combine.MCMC(DP_mcmc2020)

CR_file2020 <- readRDS("fit_CR2020_idx523_official.rds")
CR_mcmc2020 <- as.mcmc.list(lapply(1:3,function(x){as.mcmc(CR_file2020[[x]]$samples)}))
combined_listCR2020 <- combine.MCMC(CR_mcmc2020)

## 2021

DP_file2021 <- readRDS("fit_DP2021_idx127_new_official.rds")
DP_mcmc2021 <- as.mcmc.list(lapply(1:3,function(x){as.mcmc(DP_file2021[[x]]$samples)}))
combined_listDP2021 <- combine.MCMC(DP_mcmc2021)

CR_file2021 <- readRDS("fit_CR2021_idx399_official.rds")
CR_mcmc2021 <- as.mcmc.list(lapply(1:3,function(x){as.mcmc(CR_file2021[[x]]$samples)}))
combined_listCR2021 <- combine.MCMC(CR_mcmc2021)

## 2022

DP_file2022 <- readRDS("fit_DP2022_idx110_new_official.rds")
DP_mcmc2022 <- as.mcmc.list(lapply(1:3,function(x){as.mcmc(DP_file2022[[x]]$samples)}))
combined_listDP2022 <- combine.MCMC(DP_mcmc2022)

CR_file2022 <- readRDS("fit_CR2022_idx409_official.rds")
CR_mcmc2022 <- as.mcmc.list(lapply(1:3,function(x){as.mcmc(CR_file2022[[x]]$samples)}))
combined_listCR2022 <- combine.MCMC(CR_mcmc2022)

```

**MCMC Diagnostics.** We computed the Monte-Carlo Standard Error (MCSE), which is a measure of the precision of the posterior distribution obtained from a MCMC algorithm. The purpose of MCSE is to provide a measure of the accuracy of the estimate of the posterior distribution, which is the distribution of the parameters of interest after taking into account the data and prior information. MCSE is calculated by estimating the standard deviation of the MCMC samples of the posterior distribution, which provides a measure of the variability in the posterior estimates due to the Monte Carlo sampling process. This value is then divided by the square root of the effective sample size (ESS), which is the number of independent samples that the MCMC algorithm generates. 

\  

```{r, echo=FALSE}
getsummary2 <- function(data1, data2, BurPer = 0){
  
  MCSE1 <- MCSE2 <- NULL
  
  for(i in 1:ncol(data1)){
    Param1 <- data1[-(1:BurPer),i]
    MCSE1[i] <- LaplacesDemon::MCSE(Param1)
  }
  
  for(i in 1:ncol(data2)){
    Param2 <- data2[-(1:BurPer),i]
    MCSE2[i] <- LaplacesDemon::MCSE(Param2)
  }
    
  df1 <- data.frame(Parameter = colnames(data1), MCSE = MCSE1)
  df2 <- data.frame(Parameter = colnames(data2), MCSE = MCSE2)
  
  
  df <- rbind(df1,df2)[-c(1:7),]
  rownames(df) <- NULL
  df
}
```

**Year 2020.** 

```{r, echo=FALSE}
MCSE_summary = getsummary2(data1 = combined_listCR2020, data2 = combined_listDP2020)
kbl(MCSE_summary, booktabs = T, longtable = T) %>%
kable_styling(latex_options = c("hold_position","striped"))
```

\newpage 

**Year 2021.** 

```{r, echo=FALSE}
MCSE_summary = getsummary2(data1 = combined_listCR2021, data2 = combined_listDP2021)
kbl(MCSE_summary, booktabs = T, longtable = T) %>%
kable_styling(latex_options = c("hold_position","striped"))
```

\newpage 

**Year 2022.** 

```{r, echo=FALSE}
MCSE_summary = getsummary2(data1 = combined_listCR2022, data2 = combined_listDP2022)
kbl(MCSE_summary, booktabs = T, longtable = T) %>%
kable_styling(latex_options = c("hold_position","striped"))
```

\ 

